---
title: "R Notebook"
output: html_notebook
---


load packages
```{r, include=FALSE}
library(tidyverse)  
library(tools)  
library(lubridate)
```



load pre-edited receiver metadata data
```{r}

#Contains columns: array (ACT code), location, lat, lon, depth_m (bottom depth), deploy_datetime, year, receiver
dep <- read.csv("example_csv_files/deps.csv")

```



load fathom CSV files and keep only detection data
```{r}

#for the sake of this example, we provide a dataframe format akin to the expected result
dets <- read.csv("example_csv_files/dets.csv")

#however, if you need to load fathom csv files use the code that follows:
parse_fathom <- function(file) {
  
  lines <- read_lines(file)
  #identify detection lines
  det_lines   <- lines[str_detect(lines, "^DET,")]
  # Parse each independently
  df_det <- if (length(det_lines) > 0)
    read_csv(paste(det_lines, collapse="\n"),
             col_names = FALSE, show_col_types = FALSE)
  else tibble()
  # Return separately
  list(det = df_det)}

#load Files
folder <- c(
  "enter location of folder containing fathom csv's")

#List all CSV files
files <- list.files(folder, pattern = "\\.csv$", full.names = TRUE)

#Parse each file
parsed <- lapply(files, parse_fathom)

#Name entries after filename 
names(parsed) <- tools::file_path_sans_ext(basename(files))
 
#Combine detections
dets <- bind_rows(
  lapply(names(parsed), function(n) {
    if (nrow(parsed[[n]]$det) > 0)
      parsed[[n]]$det |> mutate(file = n)}), .id = NULL)

dets <- dets %>%
 dplyr::mutate(receiver = paste0(X6,"-",X7),
               transmitter = X10,
               datetime = ymd_hms(X2)) %>% 
 dplyr::select(receiver,datetime,transmitter) %>%
 dplyr::mutate(year = year(datetime)) %>%
 unique()

  #in case you need it for VR2Tx or : 
dets$receiver <- gsub("-69-", "-", dets$receiver)

#resulting dataframe has columns: receiver, datetime (at detection), year (of detection), and transmitter)

#remove unecessary items from environment
rm(parse_fathom,parsed,files,folder)

```



merge receiver metadata and detection data
```{r}

a_master <- merge(dets,
               dep,
               by = c(
                 "receiver",
                 "year"))

#clean to leave only what you want
a_master <- a_master %>%
  dplyr::select(
    location,
    receiver,
    lat,
    lon,
    year,
    transmitter,
    datetime,
    depth_m) %>%
  dplyr::distinct()

```




summarize sites for shinyapp 
```{r}

data.deps <- dep %>%
  dplyr::group_by(location) %>%
  dplyr::mutate(yrs_avail = n_distinct(year)) %>%
  dplyr::ungroup() %>%
  #toggle the filter below depending on your planned reporting period. Depending on the year you are reporting, or if you are reporting all years, you will need to edit the code that follows.  
  dplyr::filter(year == 2023) %>%
  dplyr::select(array, location, lat, lon, depth_m, deploy_datetime, yrs_avail) %>%
  #extract one row per location (the most recent deployment)
  dplyr::group_by(location) %>%
  dplyr::slice_max(deploy_datetime, n = 1, with_ties = FALSE) %>%
  dplyr::ungroup() %>%
  #this arranges the sites for assigning number ID's
  dplyr::arrange(lon, lat) %>%
  dplyr::mutate(
    ID = dplyr::row_number(),
    depth_m = paste0(round(depth_m, 0), " m"),
    deploy_date = as.Date(deploy_datetime),
    #below, you enter the date/month that receivers were last checked.
    lastCheck = "December 2023", 
    lastUpdated = Sys.Date(),
    #below, you enter when they will be checked next
    nextCheck = "Spring 2026",
    #enter sponsoring body
    sponsor = dplyr::case_when(
      array == "NGOMWS" ~ "Maine"),
    #when Rx site was established
    origin = dplyr::case_when(
      array == "NGOMWS" ~ 2021),
    #any info you want to display
    aboutSite = dplyr::case_when(
      array == "NGOMWS" ~ "One of several sites used in the Maine acoustics program")) %>%     
  dplyr::select(ID, location, lat, lon, depth_m, deploy_date, lastCheck, 
                lastUpdated, yrs_avail, sponsor, origin, nextCheck, aboutSite)


```



get count of unique transmitters that have been detected 
```{r}

data.dets <- a_master %>%
  group_by(location) %>%
  summarise(
    
    totalTransmitters = n_distinct(transmitter),
    transmittersDet2023 = {
      t2023 <- unique(transmitter[year == 2023])
      if(length(t2023) > 0) {
        paste0(length(t2023), ": ", paste(t2023, collapse = ", "))
      } else {
        "0"}
    },
    .groups = "drop")

```



calculate "visits" from your transmitters
```{r}

visits <- a_master %>%
  dplyr::arrange(location, transmitter, datetime) %>%
  dplyr::group_by(location, transmitter) %>%
  dplyr::mutate(
    time_diff = as.numeric(difftime(datetime, dplyr::lag(datetime), units = "mins")),
    is_new_event = dplyr::if_else(is.na(time_diff) | time_diff > 60, 1, 0),
    event_id = cumsum(is_new_event)) %>%
  dplyr::group_by(location, transmitter, event_id) %>%
  dplyr::summarise(firstdatetime = min(datetime), .groups = "drop")

data.visit <- visits %>%
  dplyr::mutate(year = lubridate::year(firstdatetime)) %>%
  dplyr::group_by(location) %>%
  dplyr::summarise(
    totalVisits = dplyr::n(),
    transmitterVisits2023 = sum(year == 2023),
    .groups = "drop")

```



get day-specific visit data
```{r}

format_dates <- function(dates) {
  
  dates <- as.Date(dates)
  
  tibble(date = dates) %>%
    mutate(
      year  = lubridate::year(date),
      month = lubridate::month(date),
      month_lbl = format(date, "%b"),
      day = lubridate::day(date)) %>%
    arrange(year, month, day) %>%
    group_by(year, month, month_lbl) %>%
    summarise(
      month_text = {
        d <- day
        
        runs <- split(d, cumsum(c(1, diff(d) != 1)))
        
        paste0(
          first(month_lbl), " ",
          paste(
            vapply(runs, function(r) {
              if (length(r) >= 3) {
                paste0(min(r), "-", max(r))
              } else {
                paste(r, collapse = ", ")
              }
            }, character(1)),
            collapse = ", "
          )
        )
      },
      .groups = "drop"
    ) %>%
    pull(month_text) %>%
    paste(collapse = "; ")
}


data.visit.days <- visits %>%
  ungroup() %>%
  dplyr::mutate(firstEST = as.Date(firstdatetime, tz = 'EST')) %>%
  filter(year(firstEST) == 2023) %>%
  dplyr::select(location,firstEST) %>%
  unique() %>%
  dplyr::group_by(location) %>%
  summarise(
    dates2023 = format_dates(firstEST),
    .groups = "drop")

rm(format_dates)

```



merge data into one final df
```{r}

data.final <- data.deps %>%
  left_join(data.dets, by = "location") %>%
  left_join(data.visit, by = "location") %>%
  left_join(data.visit.days, by = "location") %>%
  # Clean up NAs
  mutate(
    transmittersDet2023 = replace_na(transmittersDet2023, "0"),
    totalTransmitters = replace_na(totalTransmitters, 0),
    transmitterVisits2023 = replace_na(transmitterVisits2023, 0),
    totalVisits = replace_na(totalVisits, 0),
    dates2023 = replace_na(dates2023, ""))

```



Check that data.final is good to go. Add in sensorLoc, which is where in the water column the receiver was moored.
```{r}
data.final <- data.final %>%
  dplyr::mutate(sensorLoc = case_when(
    ID %in% 1:7 ~ "surface"
  ))

data.final$dates2023[is.na(data.final$dates2023)] <- ""
```



Export into your data subfolder within the shinyapp folder
```{r}

output_path <- "../2_shinyApp/data/data.csv"
write.csv(data.final, file = output_path, row.names = FALSE)

```
