---
title: "R Notebook"
output: html_notebook
---


load packages
```{r, warning = F}
library(tidyverse)
library(dplyr)
library(plotly)
```



load the resulting data from 1_mapData and site metadata
```{r}

output_path <- "../2_shinyApp/data/data.csv"
data <- read.csv(file = output_path)

#Contains columns: array (ACT code), location, lat, lon, depth_m (bottom depth), deploy_datetime, year, receiver
dep <- read.csv("example_csv_files/deps.csv")

```



load fathom CSV files and keep only temperature data
```{r}

#for the sake of this example, we provide a dataframe format akin to the expected result
temps <- read.csv("example_csv_files/temps.csv")

#however, if you need to load fathom csv files use the code that follows:
parse_fathom <- function(file) {
  
  lines <- read_lines(file)
  #identify detection lines
  temp_lines   <- lines[str_detect(lines, "^TEMP,")]
  # Parse each independently
  df_temp <- if (length(temp_lines) > 0)
    read_csv(paste(temp_lines, collapse="\n"),
             col_names = FALSE, show_col_types = FALSE)
  else tibble()
  # Return separately
  list(temp = df_temp)}

#load Files
folder <- c(
  "enter location of folder containing fathom csv's")

#List all CSV files
files <- list.files(folder, pattern = "\\.csv$", full.names = TRUE)

#Parse each file
parsed <- lapply(files, parse_fathom)

#Name entries after filename 
names(parsed) <- tools::file_path_sans_ext(basename(files))
 
#Combine detections
temps <- bind_rows(
  lapply(names(parsed), function(n) {
    if (nrow(parsed[[n]]$temp) > 0)
      parsed[[n]]$temp |> mutate(file = n)}), .id = NULL)

temps <- temps %>%
 dplyr::mutate(receiver = paste0(X6,"-",X7),
               receiver = gsub("-69-", "-", receiver),
               temp = X8,
               datetime = X2) %>%
  dplyr::select(receiver,datetime,temp) %>%
  unique()

rm(parsed,files,folder,parse_fathom)

```



assign ID to temp data,extract date
```{r}

temps <- temps %>%
  left_join(dep %>% select(receiver, location), by = "receiver") %>%
  unique()

temps <- merge(
  temps,
  data,
  by = "location") %>%
  dplyr::select(ID,location,receiver,datetime,temp) %>%
  dplyr::mutate(date = as.Date(datetime)) %>%
  unique()

```



Here we are going to filter out temp data that may not be useful to us (i.e., before or after receiver was deployed) and switch from C to F. Use a plot in this first chunk to view data, and then filter in the second chunk. 
```{r}
plot_data <- temps %>%
  arrange(ID, datetime)
receiver_Names <- sort(unique(plot_data$ID))

plot_all <- plot_ly()

for (i in seq_along(receiver_Names)) {
  r <- receiver_Names[i]
  df <- plot_data %>% filter(ID == r)
  plot_all <- plot_all %>%
    add_lines(
      data = df,
      x = ~datetime,
      y = ~temp,
      name = paste("Receiver", r),
      visible = ifelse(i == 1, TRUE, FALSE))}

plot_all <- plot_all %>%
  layout(
    title = "Receiver Temperature Inspection",
    xaxis = list(title = "Date"),
    yaxis = list(title = "Temperature"),
    updatemenus = list(
      list(
        type = "dropdown",
        active = 0,
        buttons = lapply(seq_along(receiver_Names), function(i) {
          list(
            method = "update",
            args = list(
              list(visible = seq_along(receiver_Names) == i),
              list(title = paste("Receiver", receiver_Names[i], "Temperature"))),
            label = receiver_Names[i])}))))

plot_all
```


```{r}
temps <- temps %>%
  filter(
    (ID = 1 & 
      (between(date, as.Date("2023-05-17"), as.Date("2023-11-14")))))  %>%
  dplyr::mutate(temp = (temp * (9/5)) + 32)
```


now make sure the resulting temp data looks right
```{r}
plot_data <- temps %>%
  group_by(ID) %>%
  arrange(datetime) %>%
  mutate(
    gap_hours = as.numeric(difftime(datetime, lag(datetime), units = "hours")),
    temp = ifelse(!is.na(gap_hours) & gap_hours > 24, NA, temp)
  ) %>%
  ungroup()
receiver_IDs <- sort(unique(plot_data$ID))

plot_all <- plot_ly()

for (i in seq_along(receiver_IDs)) {
  r <- receiver_IDs[i]
  df <- plot_data %>% filter(ID == r)

  plot_all <- plot_all %>%
    add_lines(
      data = df,
      x = ~datetime,
      y = ~temp,
      name = paste("Receiver", r),
      visible = ifelse(i == 1, TRUE, FALSE)
    )
}

plot_all <- plot_all %>%
  layout(
    title = "Receiver Temperature Inspection",
    xaxis = list(title = "Date"),
    yaxis = list(title = "Temperature"),
    updatemenus = list(
      list(
        type = "dropdown",
        active = 0,
        buttons = lapply(seq_along(receiver_IDs), function(i) {
          list(
            method = "update",
            args = list(
              list(visible = seq_along(receiver_IDs) == i),
              list(title = paste("Receiver", receiver_IDs[i], "Temperature"))
            ),
            label = receiver_IDs[i]
          )
        })
      )
    )
  )

plot_all
```


export 
```{r}

output_path <- "../2_shinyApp/data/temp.csv"
write.csv(temps, file = output_path, row.names = FALSE)

```